Visualization and analysis of probability distributions of large temporal data

### 2021-11-13 Rmd/05-conclusion.Rmd
another go at conclusion and future work

The **gravitas** package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities. The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together.) This package was developed as part of my internship at Google Summer of Code, 2019. It has been on CRAN since January 2020. The website (<https://sayani07.github.io/gravitas>) includes full documentation and two vignettes about the package usage. There has been a total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01. This package supplements the paper corresponding to Chapter \@ref(ch:gravitas), which has won the ACEMS Business Analytics Award 2021. The package can be generalized to non-temporal applications for which a hierarchical structure can be construed similar to time.

@@

The R package hakear (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch:hakear). The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$. These selected harmonies can be plotted using package `gravitas` for potentially interesting displays. This package is reliant on parallel processing using multiple multi-core computers for faster computation of $wpd$.

@@

#### Non-stationary time series

@@

#### Structured missingness

It is possible that data for a whole category of cyclic granularity is unavailable or that there are insufficient observations to compute distributions. For example, a customer does not have data for a particular day of the week or month throughout their observation period. While visualizing probability distributions across categories in Chapters \@ref(ch:gravitas) and \@ref(ch:hakear), this can be indicated by displaying dot plots instead of summarizing distributions. But the distances in Chapter \@ref(ch:gracsr) can not handle missing observations if they are structured like this. The distances have to be designed to incorporate customers with structural missingness and more research is needed to fully comprehend its implications while visually characterizing them.

#### Anomalies

Characterizing clusters with varied or outlying customers can result in patterns that do not represent the group. Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful. An appropriate anomaly detection approach could be applied to filter the anomalies, and then a decision has to be made regarding if the monsters should form a group or be assigned to the closest group through a classification method.

### Develop inferential methods for the clusters

@@

#### Scaling up for larger uncertainty and computational efficiency

@@

### Scaling up the methods for multivariate time series data

Another possible extension would be to create a similar framework for visualizing and analyzing multivariate time series data. With multiple time series available for each observation, the complexity of efficient exploration and visualization grows exponentially. In this case, conditional distributions include not only temporal dependency but also variables and their dependencies. This adds to the already high-dimensional data structures that result from studying distributions. This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing one or a few multivariate time series data. Then a similar approach of unsupervised clustering can be used to group multiple time series over multiple time granularities and multiple variables. This is a similar approach to that followed in this thesis for handling univariate time series.

>>>>

The `gravitas` package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities. The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together.) This package was developed as part of my internship at Google Summer of Code, 2019. It has been on CRAN since January 2020. The website (<https://sayani07.github.io/gravitas>) includes full documentation and two vignettes about the package usage. There has been a total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01. This package supplements the paper corresponding to Chapter \@ref(ch:gravitas), which has won the ACEMS Business Analytics Award 2021. The package can be generalized to non-temporal applications for which a hierarchical structure can be construed similar to time.

@@

The R package `hakear` (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch:hakear). The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$. These selected harmonies can be plotted using package `gravitas` for potentially interesting displays. This package is reliant on parallel processing using multiple multi-core computers for faster computation of $wpd$.

@@

**Non-stationary time series**

@@

**Structured missingness**

It is possible that data for a whole category of cyclic granularity is unavailable or that there are insufficient observations to compute distributions. For example, a customer does not have data for a particular day of the week or month throughout their observation period. While visualizing probability distributions across categories in Chapters \@ref(ch:gravitas) and \@ref(ch:hakear), this can be indicated by displaying dot plots instead of summarizing distributions. But the distances in Chapter \@ref(ch:gracsr) can not handle missing observations if they are structured like this. More research is needed to design a distance that can incorporate customers with structural missingness and also comprehend its implications while visually characterizing them.

**Anomalies**

Characterizing clusters with varied or outlying customers can result in patterns that do not represent the group. Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful. An appropriate anomaly detection approach could be applied to filter the anomalies, and then a decision has to be made regarding if the anomalies should form a group or be assigned to the closest group through a classification method.

**Develop inferential methods for the clusters**

@@

**Scaling up for larger uncertainty and computational efficiency**

@@

**Scaling up the methods for multivariate time series data**

Another possible extension would be to create a similar framework for visualizing and analyzing multivariate time series data. With multiple time series available for each observation, the complexity of efficient exploration and visualization grows exponentially. In this case, conditional distributions include not only temporal dependency but also variables and their dependencies. This adds to the already high-dimensional data structures that result from studying distributions. This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing one or a few multivariate time series data. Unsupervised clustering can then be used to group multiple time series across multiple time granularities and variables. This is a method similar to the one used in this thesis for dealing with univariate time series.
